\documentclass{article}

\usepackage{tikz} 
\usetikzlibrary{automata, positioning, arrows.meta} 

\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{fullpage}
\usepackage{color}
\usepackage{parskip}
\usepackage{hyperref}
  \hypersetup{
    colorlinks = true,
    urlcolor = blue,       % color of external links using \href
    linkcolor= blue,       % color of internal links 
    citecolor= blue,       % color of links to bibliography
    filecolor= blue,        % color of file links
    }
    
\usepackage{listings}
\usepackage[utf8]{inputenc}                                                    
\usepackage[T1]{fontenc}                                                       

\definecolor{dkgreen}{rgb}{0,0.6,0}
\definecolor{gray}{rgb}{0.5,0.5,0.5}
\definecolor{mauve}{rgb}{0.58,0,0.82}

\lstset{frame=tb,
  language=haskell,
  aboveskip=3mm,
  belowskip=3mm,
  showstringspaces=false,
  columns=flexible,
  basicstyle={\small\ttfamily},
  numbers=none,
  numberstyle=\tiny\color{gray},
  keywordstyle=\color{blue},
  commentstyle=\color{dkgreen},
  stringstyle=\color{mauve},
  breaklines=true,
  breakatwhitespace=true,
  tabsize=3
}

\usepackage{booktabs}

\newtheoremstyle{theorem}
  {\topsep}   % ABOVESPACE
  {\topsep}   % BELOWSPACE
  {\itshape\/}  % BODYFONT
  {0pt}       % INDENT (empty value is the same as 0pt)
  {\bfseries} % HEADFONT
  {.}         % HEADPUNCT
  {5pt plus 1pt minus 1pt} % HEADSPACE
  {}          % CUSTOM-HEAD-SPEC
\theoremstyle{theorem} 
   \newtheorem{theorem}{Theorem}[section]
   \newtheorem{corollary}[theorem]{Corollary}
   \newtheorem{lemma}[theorem]{Lemma}
   \newtheorem{proposition}[theorem]{Proposition}
\theoremstyle{definition}
   \newtheorem{definition}[theorem]{Definition}
   \newtheorem{example}[theorem]{Example}
\theoremstyle{remark}    
  \newtheorem{remark}[theorem]{Remark}

\title{CPSC-406 Report}
\author{Savana Chou  \\ Chapman University}

\date{\today} 

\begin{document}

\maketitle

\begin{abstract}
\end{abstract}

\setcounter{tocdepth}{3}
\tableofcontents

\section{Introduction}\label{intro}

\section{Week by Week}\label{homework}

\subsection{Week 1}

\section*{Exercise 1}

\subsection*{1. Word acceptance table}

\begin{center}
\begin{tabular}{c|c|c}
$w$ & accepted by $A_1$? & accepted by $A_2$? \\
\hline
$aaa$ & No & Yes \\
$aab$ & Yes & No \\
$aba$ & No & No \\
$abb$ & No & No \\
$baa$ & No & Yes \\
$bab$ & No & No \\
$bba$ & No & No \\
$bbb$ & No & No \\
\end{tabular}
\end{center}

\subsection*{2. Description of the languages}

\paragraph{Language of $A_1$.}

In $A_1$, any word beginning with $b$ immediately goes to a trap state.
If the word begins with $a$, the automaton stays in state $2$ while
reading $a$'s, and moves to the accepting state $3$ upon reading a
single $b$. Any further symbol leaves the accepting state.

Therefore, a word is accepted if and only if it consists of one or more
$a$'s followed by exactly one $b$.

\[
L(A_1) = \{ a^n b \mid n \ge 1 \}
\]

Equivalently, as a regular expression:

\[
L(A_1) = a^+ b
\]

\paragraph{Language of $A_2$.}

In $A_2$, the only accepting state is $3$. The automaton reaches state
$3$ precisely after reading two consecutive $a$'s. Any $b$ resets the
automaton to state $1$.

Thus, a word is accepted if and only if it ends with at least two
consecutive $a$'s.

\[
L(A_2) = \{ w \in \{a,b\}^* \mid w \text{ ends with } aa \}
\]

Equivalently, as a regular expression:

\[
L(A_2) = (a|b)^* aa
\]

\section*{Exercise 2}

Alphabet: $\Sigma = \{a,b\}$.

\subsection*{1. DFAs for the given languages}

\paragraph{(1) All words that end with \texttt{ab}.}

Let $M_1 = (Q,\Sigma,\delta,q_0,F)$ where

\[
Q = \{q_0,q_1,q_2\}, \quad q_0 \text{ start}, \quad F = \{q_2\}.
\]

Transitions:

\[
\begin{array}{c|cc}
 & a & b \\
\hline
q_0 & q_1 & q_0 \\
q_1 & q_1 & q_2 \\
q_2 & q_1 & q_0
\end{array}
\]

State $q_2$ represents that the word currently ends in \texttt{ab}.

\bigskip

\paragraph{(2) All words that contain \texttt{aba}.}

Let $M_2 = (Q,\Sigma,\delta,q_0,F)$ where

\[
Q = \{q_0,q_1,q_2,q_3\}, \quad q_0 \text{ start}, \quad F = \{q_3\}.
\]

Transitions:

\[
\begin{array}{c|cc}
 & a & b \\
\hline
q_0 & q_1 & q_0 \\
q_1 & q_1 & q_2 \\
q_2 & q_3 & q_0 \\
q_3 & q_3 & q_3
\end{array}
\]

State $q_3$ means the substring \texttt{aba} has been seen.

\bigskip

\paragraph{(3) Odd number of \texttt{a}'s and odd number of \texttt{b}'s.}

Use four states representing parity:

\[
Q = \{(E,E),(E,O),(O,E),(O,O)\}
\]

(start state $(E,E)$).  
Accepting state: $(O,O)$.

Reading $a$ toggles the first component;  
reading $b$ toggles the second component.

\bigskip

\paragraph{(4) Even number of \texttt{a}'s and odd number of \texttt{b}'s.}

Same construction as (3).

Start state: $(E,E)$.  
Accepting state: $(E,O)$.

Transitions again toggle parity accordingly.

\bigskip

\paragraph{(5) Any three consecutive characters contain at least one \texttt{a}.}

This is equivalent to saying the word does \emph{not} contain \texttt{bbb}.

Use states counting consecutive $b$'s:

\[
Q = \{q_0,q_1,q_2,q_3\}
\]

\begin{itemize}
\item $q_0$: no recent $b$
\item $q_1$: one consecutive $b$
\item $q_2$: two consecutive $b$'s
\item $q_3$: three consecutive $b$'s (trap)
\end{itemize}

Start: $q_0$.

Accepting states: $\{q_0,q_1,q_2\}$.

Transitions:

- On $a$: go to $q_0$ from any state except $q_3$
- On $b$:
  \[
  q_0 \to q_1,\quad
  q_1 \to q_2,\quad
  q_2 \to q_3,\quad
  q_3 \to q_3
  \]

\bigskip

\paragraph{(6) All words that contain \texttt{bbb}.}

Same structure as (5), but now:

Accepting state: $\{q_3\}$ only.

Once $q_3$ is reached, stay there.

\bigskip

\subsection*{2. Observations}

Several patterns appear:

\begin{itemize}
\item Languages involving substrings (e.g.\ \texttt{aba}, \texttt{bbb})
      use states that track partial progress toward matching the substring.
\item Parity conditions use a product construction (Cartesian product)
      of two 2-state automata, giving 4 states.
\item Conditions about avoiding a pattern (problem 5) are closely related
      to conditions about containing that pattern (problem 6);
      they differ mainly in which states are accepting.
\item Many DFAs follow a systematic design pattern:
      track exactly the minimal information necessary to decide acceptance.
\end{itemize}

\subsection*{Week 1 Discord Question}
Why is it important that the transition function of a DFA be total (defined for every state-symbol pair)? What would break, both formally and conceptually, if we allowed “missing” transitions?

\subsection{Week 2}

\section*{Exercise 1: Product Automata}

Let $\Sigma = \{a,b\}$.

\subsection*{1. Description of $L(\mathcal A^{(1)})$ and $L(\mathcal A^{(2)})$}

\subsubsection*{Language of $\mathcal A^{(1)}$}

In $\mathcal A^{(1)}$, states $2$ and $4$ are accepting, and state $3$ is a trap state with self-loops on both $a$ and $b$. 
From the transition structure we observe:

\begin{itemize}
    \item From state $2$, reading $a$ leads to the trap state.
    \item From state $4$, reading $b$ leads to the trap state.
\end{itemize}

Thus any occurrence of the substring $aa$ or $bb$ forces the automaton into the trap state.

Therefore,
\[
L(\mathcal A^{(1)}) 
= \{ w \in \{a,b\}^* \mid w \text{ contains no } aa \text{ and no } bb \}.
\]

Equivalently, $L(\mathcal A^{(1)})$ consists of all strings whose symbols strictly alternate.

\bigskip

\subsubsection*{Language of $\mathcal A^{(2)}$}

In $\mathcal A^{(2)}$, state $2$ is the only accepting state and state $3$ is a trap.

\begin{itemize}
    \item From the start state $1$, reading $b$ leads immediately to the trap.
    \item Thus every accepted string must begin with $a$.
    \item The automaton alternates between states $1$ and $2$ on every input symbol.
\end{itemize}

Since state $2$ is accepting, the automaton accepts precisely those strings that:
\begin{itemize}
    \item start with $a$, and
    \item have odd length.
\end{itemize}

Hence,
\[
L(\mathcal A^{(2)}) 
= \{ w \in \{a,b\}^* \mid w \text{ starts with } a \text{ and } |w| \text{ is odd} \}.
\]

\bigskip

\subsection*{2. Construction of the Intersection Automaton}
\tikzset{
  automaton/.style={
    >=stealth,
    shorten >=1pt,
    node distance=2.0cm,
    every state/.style={thick, minimum size=7mm},
    initial text={},
  }
}

The product automaton has state set
\[
Q = Q^{(1)} \times Q^{(2)} = \{1,2,3,4\} \times \{1,2,3\}, \quad \text{start state } (1,1),
\]
with transition function defined by
\[
\delta\big((i,j), x\big) = \big(\delta_1(i,x), \delta_2(j,x)\big).
\]

Reachable states from $(1,1)$ and their transitions:

\begin{center}
\begin{tabular}{c|cc}
State & on $a$ & on $b$ \\ \hline
$(1,1)$ & $(2,2)$ & $(4,3)$ \\
$(2,2)$ & $(3,1)$ & $(4,1)$ \\
$(4,3)$ & $(3,3)$ & $(2,3)$ \\
$(3,1)$ & $(3,2)$ & $(3,3)$ \\
$(4,1)$ & $(3,2)$ & $(2,3)$ \\
$(3,3)$ & $(3,3)$ & $(3,3)$ \\
$(2,3)$ & $(3,3)$ & $(4,3)$ \\
$(3,2)$ & $(3,1)$ & $(3,1)$
\end{tabular}
\end{center}

Thus, the reachable states are
\[
\{(1,1), (2,2), (4,3), (3,1), (4,1), (3,3), (2,3), (3,2)\}.
\]

The accepting states of the intersection automaton are
\[
F = F^{(1)} \times F^{(2)} = \{2,4\} \times \{2\} = \{(2,2), (4,2)\}.
\]
Since $(4,2)$ is unreachable, the only accepting state is $(2,2)$.

\medskip
\begin{center}
\begin{tikzpicture}[automaton, node distance=2.2cm]
\node[state, initial] (11) {$(1,1)$};
\node[state, accepting, right=of 11] (22) {$(2,2)$};
\node[state, below=of 11] (43) {$(4,3)$};
\node[state, right=of 22] (31) {$(3,1)$};
\node[state, right=of 43] (41) {$(4,1)$};
\node[state, below right=1.5cm and 1cm of 41] (33) {$(3,3)$};
\node[state, below=of 41] (23) {$(2,3)$};
\node[state, above right=1cm and 1cm of 31] (32) {$(3,2)$};

\path[->]
  (11) edge node[above]{$a$} (22)
  (11) edge node[left]{$b$} (43)
  (22) edge node[above]{$a$} (31)
  (22) edge[bend left=15] node[below left]{$b$} (41)
  (43) edge[bend left=20] node[below]{$a$} (33)
  (43) edge node[right]{$b$} (23)
  (31) edge node[above right]{$a$} (32)
  (31) edge node[below right]{$b$} (33)
  (41) edge[bend right=25] node[left]{$a$} (32)
  (41) edge node[above]{$b$} (23)
  (33) edge[loop right] node{$a,b$} ()
  (23) edge node[above right]{$a$} (33)
  (23) edge[bend left=15] node[left]{$b$} (43)
  (32) edge[bend right=15] node[right]{$a,b$} (31);
\end{tikzpicture}
\end{center}

\subsection*{3. Correctness of the Product Construction}

By definition of the product construction, a string $w$ is accepted by $\mathcal A$ if and only if:

\begin{itemize}
    \item the run of $w$ in $\mathcal A^{(1)}$ ends in an accepting state, and
    \item the run of $w$ in $\mathcal A^{(2)}$ ends in an accepting state.
\end{itemize}

Thus,
\[
L(\mathcal A)
=
L(\mathcal A^{(1)}) \cap L(\mathcal A^{(2)}).
\]

\bigskip

\subsection*{4. Construction for the Union}

To obtain an automaton $\mathcal A'$ such that
\[
L(\mathcal A') = L(\mathcal A^{(1)}) \cup L(\mathcal A^{(2)}),
\]
we keep:

\begin{itemize}
    \item the same state set $Q_1 \times Q_2$,
    \item the same start state $(1,1)$,
    \item the same transition function.
\end{itemize}

We change only the accepting states to:

\[
F' = (F_1 \times Q_2) \cup (Q_1 \times F_2).
\]

That is, a state $(p,q)$ is accepting if either
\[
p \in F_1 \quad \text{or} \quad q \in F_2.
\]

\section*{Exercise 2: More Automata}

Let $\Sigma = \{a,b\}$.

\subsection*{1. Description of the Languages}

\subsubsection*{Language of $\mathcal B^{(1)}$}

The automaton $\mathcal B^{(1)}$ has states 
$\{p_0,p_1,p_2\}$ where $p_0$ is the start and only accepting state.
Each $b$ leaves the state unchanged, while each $a$ moves cyclically:
\[
p_0 \xrightarrow{a} p_1 \xrightarrow{a} p_2 \xrightarrow{a} p_0.
\]

Thus the automaton counts the number of $a$'s modulo $3$. Since only
$p_0$ is accepting, we obtain
\[
L(\mathcal B^{(1)}) 
= 
\{\, w \in \{a,b\}^* \mid \#_a(w) \equiv 0 \pmod{3} \,\}.
\]

In words: all strings whose number of $a$'s is divisible by $3$.

\bigskip

\subsubsection*{Language of $\mathcal B^{(2)}$}

The automaton $\mathcal B^{(2)}$ has states 
$\{q_0,q_1,q_2\}$ where $q_0$ is the start and only accepting state,
and $q_2$ is a trap state with loops on both $a$ and $b$.

The transitions show:
\[
q_0 \xrightarrow{a} q_1, 
\qquad
q_1 \xrightarrow{a} q_2.
\]

Thus any occurrence of two consecutive $a$'s sends the automaton
to the trap state $q_2$. Therefore,
\[
L(\mathcal B^{(2)}) 
=
\{\, w \in \{a,b\}^* \mid w \text{ does not contain } aa \,\}.
\]

\bigskip

\subsection*{2. Construction of the Intersection Automaton}

Define the state set as the Cartesian product:
\[
Q = \{p_0, p_1, p_2\} \times \{q_0, q_1, q_2\}, \quad \text{with start state } (p_0, q_0).
\]

We compute the reachable states and transitions:

\begin{center}
\begin{tabular}{c|cc}
State & input $a$ & input $b$ \\ \hline
$(p_0,q_0)$ & $(p_1,q_1)$ & $(p_0,q_0)$ \\
$(p_1,q_1)$ & $(p_2,q_2)$ & $(p_1,q_0)$ \\
$(p_2,q_2)$ & $(p_0,q_2)$ & $(p_2,q_2)$ \\
$(p_1,q_0)$ & $(p_2,q_1)$ & $(p_1,q_0)$ \\
$(p_0,q_2)$ & $(p_1,q_2)$ & $(p_0,q_2)$ \\
$(p_2,q_1)$ & $(p_0,q_2)$ & $(p_2,q_0)$ \\
$(p_1,q_2)$ & $(p_2,q_2)$ & $(p_1,q_2)$ \\
$(p_2,q_0)$ & $(p_0,q_1)$ & $(p_2,q_0)$ \\
$(p_0,q_1)$ & $(p_1,q_2)$ & $(p_0,q_0)$
\end{tabular}
\end{center}

The accepting states of the intersection automaton are the product of the accepting states of the original DFAs:
\[
F = F^{(1)} \times F^{(2)} = \{p_0, p_2\} \times \{q_0\} = \{(p_0,q_0), (p_2,q_0)\}.
\]

Among the reachable states, only $(p_0,q_0)$ and $(p_2,q_0)$ satisfy this condition.

\begin{center}
\begin{tikzpicture}[automaton, node distance=2cm]
\node[state, accepting, initial] (p0q0) {$(p_0,q_0)$};
\node[state, right=of p0q0] (p1q1) {$(p_1,q_1)$};
\node[state, below=of p1q1] (p1q0) {$(p_1,q_0)$};
\node[state, accepting, below=of p1q0] (p2q0) {$(p_2,q_0)$};
\node[state, right=of p1q1] (p2q2) {$(p_2,q_2)$};
\node[state, right=of p1q0] (p2q1) {$(p_2,q_1)$};
\node[state, below=of p2q2] (trap) {$(p_i,q_2)$};

\path[->]
  (p0q0) edge[loop above] node{$b$} ()
  (p0q0) edge node[above]{$a$} (p1q1)
  (p1q1) edge node[right]{$b$} (p1q0)
  (p1q1) edge node[above]{$a$} (p2q2)
  (p1q0) edge[loop left] node{$b$} ()
  (p1q0) edge node[right]{$a$} (p2q1)
  (p2q1) edge node[below]{$b$} (p2q0)
  (p2q0) edge[loop below] node{$b$} ()
  (p2q0) edge[bend left=25] node[left]{$a$} (p0q0)
  (p2q2) edge[loop right] node{$b$} ()
  (p2q2) edge node[right]{$a$} (trap)
  (trap) edge[loop right] node{$a,b$} ();
\end{tikzpicture}
\end{center}

\noindent
\textit{Note:} The states $(p_0,q_2)$, $(p_1,q_2)$, and $(p_2,q_2)$ are merged into a single trap state for simplicity.

\paragraph{Part 3: Verification.}

By construction, a string $w$ is accepted by the intersection automaton $\mathcal{B}$ if and only if it is accepted by both original automata, i.e.,
\[
w \in L(\mathcal{B}) \iff w \in L(\mathcal{B}^{(1)}) \text{ and } w \in L(\mathcal{B}^{(2)}).
\]

Hence, the language recognized by $\mathcal{B}$ is
\[
L(\mathcal{B}) = L(\mathcal{B}^{(1)}) \cap L(\mathcal{B}^{(2)}) 
= \{w \mid \#_a(w) \not\equiv 1 \pmod{3}\} \cap \{w \mid w \text{ contains no } aa\}.
\]

\subsection*{3. Correctness of the Construction}

By definition of the product construction, a string $w$ is accepted by
$\mathcal B$ if and only if:

\begin{itemize}
    \item the run of $w$ in $\mathcal B^{(1)}$ ends in $p_0$, and
    \item the run of $w$ in $\mathcal B^{(2)}$ ends in $q_0$.
\end{itemize}

Hence,
\[
L(\mathcal B)
=
L(\mathcal B^{(1)})
\cap
L(\mathcal B^{(2)}).
\]

\bigskip

\subsection*{4. Construction of $\mathcal B'$ for the Union with a Complement}

We want an automaton $\mathcal B'$ such that
\[
L(\mathcal B')
=
L(\mathcal B^{(1)})
\cup
\overline{L(\mathcal B^{(2)})}.
\]

\paragraph{Step 1: Complement $\mathcal B^{(2)}$.}

Since $\mathcal B^{(2)}$ is a complete DFA, we obtain
$\overline{\mathcal B^{(2)}}$ by swapping accepting and non-accepting states.
Thus $q_1$ and $q_2$ become accepting, and $q_0$ becomes rejecting.

\paragraph{Step 2: Product Construction for the Union.}

We keep:

\begin{itemize}
    \item the same state set $Q_1 \times Q_2$,
    \item the same start state $(p_0,q_0)$,
    \item the same transition function.
\end{itemize}

We change the accepting states to:
\[
F'
=
(F_1 \times Q_2)
\cup
(Q_1 \times F_{\overline{2}}).
\]

That is, a state $(p,q)$ is accepting if either
\[
p = p_0
\quad \text{or} \quad
q \in \{q_1,q_2\}.
\]

\subsubsection*{Exercise 2.2.7 — State with Self-Loops}

\paragraph{Problem.}
Consider a DFA $\mathcal{A}$ and a state $q$ such that for every input symbol $a \in \Sigma$, the transition function satisfies $\delta(q, a) = q$. Prove, using induction on the length of input strings, that for any string $w \in \Sigma^*$, the extended transition function satisfies $\hat{\delta}(q, w) = q$.

\paragraph{Solution.}

We proceed by induction on the length of the string $w$, denoted $|w|$.

\medskip
\noindent\textbf{Base case:} $|w| = 0$, i.e., $w = \varepsilon$ (the empty string).  

By the definition of the extended transition function:
\[
\hat{\delta}(q, \varepsilon) = q.
\]
Hence, the statement is true for the base case.

\medskip
\noindent\textbf{Inductive step:}  

\textit{Inductive hypothesis (IH):} Assume that for all strings $w'$ of length $n$, we have $\hat{\delta}(q, w') = q$.  

\textit{Goal:} Show that for any string $w$ of length $n+1$, $\hat{\delta}(q, w) = q$.  

Let $w$ be a string of length $n+1$, and write $w = w'a$ where $|w'| = n$ and $a \in \Sigma$.  

By the recursive definition of $\hat{\delta}$:
\[
\hat{\delta}(q, w) = \hat{\delta}(q, w'a) = \delta\big(\hat{\delta}(q, w'), a\big).
\]

Applying the inductive hypothesis, $\hat{\delta}(q, w') = q$, so:
\[
\delta\big(\hat{\delta}(q, w'), a\big) = \delta(q, a).
\]

Since $q$ has a self-loop for every symbol, $\delta(q, a) = q$ for all $a \in \Sigma$. Therefore:
\[
\hat{\delta}(q, w) = q.
\]

This completes the inductive step, and by induction, the result holds for all $w \in \Sigma^*$.

\subsection*{Week 2 Discord Question}
In both exercises, we constructed product automata to recognize intersections and modified acceptance conditions to recognize unions or complements. Suppose instead we were given only one of the product automata (without being told it was constructed from two smaller DFAs). How could we determine whether this automaton can be decomposed into a product of two smaller DFAs?

\subsection{Week 3}

\subsubsection*{Homework 1}

Consider the NFA $\mathcal{A}$ over the alphabet $\Sigma = \{a, b, c\}$.

\paragraph{Part 1: Description of the language.}

By inspecting the diagram, we see that $\mathcal{A}$ accepts strings that:
\begin{itemize}
\item Begin with any combination of $a$'s and $b$'s (remaining in state $0$),
\item Followed by exactly one $a$ (moving to state $1$),
\item And end with either a $b$ (reaching accepting state $2$) or a $c$ (reaching accepting state $3$).
\end{itemize}

\[
\boxed{L(\mathcal{A}) = (a \mid b)^* a (b \mid c)}
\]

\paragraph{Part 2: Formal definition of the NFA.}

The automaton is defined as $\mathcal{A} = (Q, \Sigma, \delta, q_0, F)$ with:
\begin{itemize}
\item $Q = \{0,1,2,3\}$
\item $\Sigma = \{a,b,c\}$
\item Start state $q_0 = 0$
\item Accepting states $F = \{2,3\}$
\item Transition function $\delta: Q \times \Sigma \to \mathcal{P}(Q)$ given by:
\end{itemize}

\begin{center}
\begin{tabular}{c|ccc}
$\delta$ & $a$ & $b$ & $c$ \\ \hline
$0$ & $\{0,1\}$ & $\{0\}$ & $\emptyset$ \\
$1$ & $\emptyset$ & $\{2\}$ & $\{3\}$ \\
$2$ & $\emptyset$ & $\emptyset$ & $\emptyset$ \\
$3$ & $\emptyset$ & $\emptyset$ & $\emptyset$
\end{tabular}
\end{center}

\paragraph{Part 3: Tracing all paths for $w = abab$.}

We examine all possible transitions through $\mathcal{A}$ for the input string $abab$:

\medskip
\noindent\textbf{Path 1:} $0 \xrightarrow{a} 0 \xrightarrow{b} 0 \xrightarrow{a} 0 \xrightarrow{b} 0$ (ends in state $0$, \textbf{not accepting})

\noindent\textbf{Path 2:} $0 \xrightarrow{a} 0 \xrightarrow{b} 0 \xrightarrow{a} 1 \xrightarrow{b} 2$ (ends in state $2$, \textbf{ACCEPTING})

\noindent\textbf{Path 3:} $0 \xrightarrow{a} 1 \xrightarrow{b} 2$ (cannot continue for remaining symbols)

\medskip
\noindent Visual representation of successful paths:

\begin{center}
\begin{tikzpicture}[
  >=stealth,
  node distance=2cm,
  every state/.style={thick, minimum size=8mm},
  initial text={},
  level/.style={sibling distance=3cm, level distance=1.5cm}
]

\node[state] (00) {$0$};
\node[state, below left=1.2cm and 1.5cm of 00] (01) {$0$};
\node[state, below right=1.2cm and 1.5cm of 00] (11) {$1$};
\node[state, below left=1.2cm and 0.8cm of 01] (02) {$0$};
\node[state, below right=1.2cm and 0.8cm of 01] (12) {$1$};
\node[state, below=1.2cm of 12] (22) {$2$};
\node[state, below=1.2cm of 02] (03) {$0$};
\node[state, below=1.2cm of 03] (04) {$0$};

\path[->]
  (00) edge node[above left]{$a$} (01)
  (00) edge node[above right]{$a$} (11)
  (01) edge node[above left]{$b$} (02)
  (01) edge node[above right]{$b$} (12)
  (12) edge node[right]{$a$} (22)
  (02) edge node[left]{$a$} (03)
  (03) edge node[left]{$b$} (04);

\node[right=0.3cm of 22, accepting] {$\checkmark$};
\node[right=0.3cm of 04] {$\times$};

\end{tikzpicture}
\end{center}

Since at least one path reaches an accepting state, $w = abab$ is \textbf{accepted}.

\paragraph{Part 4: Determinization via the power set construction.}

The DFA $\mathcal{A}^D = (Q^D, \Sigma, \delta^D, q_0^D, F^D)$ is defined as:
\begin{itemize}
\item $Q^D = \mathcal{P}(Q)$ (we list only reachable subsets)
\item $q_0^D = \{0\}$
\item $F^D = \{S \subseteq Q \mid S \cap \{2,3\} \neq \emptyset\}$
\item $\delta^D(S,x) = \bigcup_{q \in S} \delta(q,x)$
\end{itemize}

\medskip
\noindent\textbf{Transition table for $\mathcal{A}^D$:}

\begin{center}
\begin{tabular}{c|ccc}
State & $a$ & $b$ & $c$ \\ \hline
$\to \{0\}$ & $\{0,1\}$ & $\{0\}$ & $\emptyset$ \\
$\{0,1\}$ & $\{0,1\}$ & $\{0,2\}$ & $\{3\}$ \\
$*\{0,2\}$ & $\{0,1\}$ & $\{0\}$ & $\emptyset$ \\
$*\{3\}$ & $\emptyset$ & $\emptyset$ & $\emptyset$ \\
$\emptyset$ & $\emptyset$ & $\emptyset$ & $\emptyset$
\end{tabular}
\end{center}

\medskip
\noindent\textbf{Transition diagram for $\mathcal{A}^D$:}

\begin{center}
\begin{tikzpicture}[
  >=stealth,
  shorten >=1pt,
  node distance=2.5cm,
  every state/.style={thick, minimum size=10mm},
  initial text={}
]

\node[state, initial] (S0) {$\{0\}$};
\node[state, right=of S0] (S01) {$\{0,1\}$};
\node[state, accepting, below=of S01] (S02) {$\{0,2\}$};
\node[state, accepting, right=of S01] (S3) {$\{3\}$};
\node[state, below=of S0] (empty) {$\emptyset$};

\path[->]
  (S0) edge[loop above] node{$b$} ()
  (S0) edge node[above]{$a$} (S01)
  (S0) edge node[left]{$c$} (empty)
  (S01) edge[loop above] node{$a$} ()
  (S01) edge node[right]{$b$} (S02)
  (S01) edge node[above]{$c$} (S3)
  (S02) edge[bend left=20] node[left]{$a$} (S01)
  (S02) edge node[below]{$b$} (S0)
  (S02) edge node[below right]{$c$} (empty)
  (S3) edge[bend right=30] node[above]{$a,b,c$} (empty)
  (empty) edge[loop below] node{$a,b,c$} ();

\end{tikzpicture}
\end{center}

The determinized automaton has \textbf{5 reachable states}.

\subsubsection*{Homework 2}

Consider the NFA $\mathcal{A}$ over $\Sigma = \{0,1\}$.

\paragraph{Formal specification.}

$\mathcal{A} = (Q, \Sigma, \delta, q_0, F)$ with:
\begin{itemize}
\item $Q = \{q_0, q_1, q_2, q_3\}$
\item Start state $q_0$
\item $F = \{q_3\}$
\item Transition function $\delta$ as follows:
\end{itemize}

\begin{center}
\begin{tabular}{c|cc}
$\delta$ & $0$ & $1$ \\ \hline
$q_0$ & $\{q_0\}$ & $\{q_0,q_1\}$ \\
$q_1$ & $\{q_2\}$ & $\emptyset$ \\
$q_2$ & $\{q_3\}$ & $\{q_1\}$ \\
$q_3$ & $\{q_3\}$ & $\{q_3\}$
\end{tabular}
\end{center}

\paragraph{Part 1: Determinization (power set construction).}

Starting from $\{q_0\}$, the reachable subsets are computed:

\begin{center}
\begin{tabular}{c|cc}
State & $0$ & $1$ \\ \hline
$\to \{q_0\}$ & $\{q_0\}$ & $\{q_0,q_1\}$ \\
$\{q_0,q_1\}$ & $\{q_0,q_2\}$ & $\{q_0,q_1\}$ \\
$\{q_0,q_2\}$ & $\{q_0,q_3\}$ & $\{q_0,q_1\}$ \\
$*\{q_0,q_3\}$ & $\{q_0,q_3\}$ & $\{q_0,q_1,q_3\}$ \\
$*\{q_0,q_1,q_3\}$ & $\{q_0,q_2,q_3\}$ & $\{q_0,q_1,q_3\}$ \\
$*\{q_0,q_2,q_3\}$ & $\{q_0,q_3\}$ & $\{q_0,q_1,q_3\}$
\end{tabular}
\end{center}

\medskip
\noindent\textbf{Transition diagram for the determinized DFA:}

\begin{center}
\begin{tikzpicture}[
  >=stealth,
  shorten >=1pt,
  node distance=2.2cm,
  every state/.style={thick, minimum size=11mm},
  initial text={}
]

\node[state, initial] (q0) {$\{q_0\}$};
\node[state, right=of q0] (q01) {$\{q_0,q_1\}$};
\node[state, below=of q01] (q02) {$\{q_0,q_2\}$};
\node[state, accepting, right=of q01] (q03) {$\{q_0,q_3\}$};
\node[state, accepting, below=of q03] (q013) {$\{q_0,q_1,q_3\}$};
\node[state, accepting, below=of q02] (q023) {$\{q_0,q_2,q_3\}$};

\path[->]
  (q0) edge[loop left] node{$0$} ()
  (q0) edge node[above]{$1$} (q01)
  (q01) edge[loop above] node{$1$} ()
  (q01) edge node[right]{$0$} (q02)
  (q02) edge node[above]{$1$} (q01)
  (q02) edge node[above]{$0$} (q03)
  (q03) edge[loop above] node{$0$} ()
  (q03) edge node[right]{$1$} (q013)
  (q013) edge[loop right] node{$1$} ()
  (q013) edge node[right]{$0$} (q023)
  (q023) edge[bend left=15] node[below]{$1$} (q013)
  (q023) edge[bend right=30] node[below left]{$0$} (q03);

\end{tikzpicture}
\end{center}

The determinized DFA has \textbf{6 reachable states}.

\paragraph{Part 2: Minimal DFA.}

Yes, a smaller DFA exists for the same language. The language consists of all strings containing the substring $100$:

\[
L(\mathcal{A}) = (0 \mid 1)^* 1 0 0 (0 \mid 1)^*
\]

\medskip
\noindent Minimal DFA with 4 states:

\begin{center}
\begin{tikzpicture}[
  >=stealth,
  shorten >=1pt,
  node distance=2.3cm,
  every state/.style={thick, minimum size=9mm},
  initial text={}
]

\node[state, initial] (s0) {$s_0$};
\node[state, right=of s0] (s1) {$s_1$};
\node[state, right=of s1] (s2) {$s_2$};
\node[state, accepting, right=of s2] (s3) {$s_3$};

\path[->]
  (s0) edge[loop above] node{$0$} ()
  (s0) edge[bend left=15] node[above]{$1$} (s1)
  (s1) edge[bend left=15] node[below]{$1$} (s0)
  (s1) edge node[above]{$0$} (s2)
  (s2) edge[bend right=40] node[below]{$1$} (s0)
  (s2) edge node[above]{$0$} (s3)
  (s3) edge[loop above] node{$0,1$} ();

\end{tikzpicture}
\end{center}

\noindent State meanings:
\begin{itemize}
\item $s_0$: no match for $100$ so far
\item $s_1$: most recently saw $1$
\item $s_2$: most recently saw $10$
\item $s_3$: matched $100$ (accepting, remains accepting)
\end{itemize}

This minimal DFA reduces the number of states from 6 to \textbf{4} by tracking progress toward the target substring.

\subsection*{Week 3 Discord Question}
Is there a structural way to predict which subsets will collapse during minimization before fully constructing the DFA? In other words, can we detect equivalent subset-states early from properties of the original NFA?

\subsection{Week 4}

\subsection*{The DFA $\mathcal{A}$}

\begin{itemize}
    \item \textbf{States:} $p = 1$, $q = 2$
    \item \textbf{Start state:} $p$ (also an accept state, indicated by double circle)
    \item \textbf{Accept states:} $\{p\}$
    \item \textbf{Transitions:}
\end{itemize}

\begin{center}
\begin{tabular}{c||c|c}
\toprule
\textbf{State} & $a$ & $b$ \\
\midrule
$\rightarrow\!*\, p$ & $p$ & $q$ \\
$q$                  & $p$ & $q$ \\
\bottomrule
\end{tabular}
\end{center}

\begin{center}
\begin{tikzpicture}[
    shorten >=2pt, node distance=3.5cm, on grid, auto,
    >={Stealth[length=6pt]}
]
    \node[state, initial, accepting] (p) {$p$};
    \node[state, right=of p]         (q) {$q$};
    \path[->]
        (p) edge[loop above]          node {$a$} (p)
        (p) edge[bend left=20]        node {$b$} (q)
        (q) edge[bend left=20]        node {$a$} (p)
        (q) edge[loop right]          node {$b$} (q);
\end{tikzpicture}
\end{center}

\section*{Kleene's Algorithm}

\subsection*{Setup}

Kleene's algorithm computes regular expressions $R^{(k)}_{ij}$ representing the set of all strings that take state $i$ to state $j$ using only intermediate states from $\{1, \ldots, k\}$.

\medskip
The recurrence is:
\[
R^{(k)}_{ij} = R^{(k-1)}_{ik}\left(R^{(k-1)}_{kk}\right)^* R^{(k-1)}_{kj} \;+\; R^{(k-1)}_{ij}
\]

with base case (no intermediate states, $k=0$):
\[
R^{(0)}_{ij} =
\begin{cases}
a_1 + a_2 + \cdots + a_m & \text{if there are transitions } i \xrightarrow{a_\ell} j \text{ (and } i \neq j\text{)} \\
a_1 + \cdots + a_m + \varepsilon & \text{if } i = j \text{ and there are transitions } i \xrightarrow{a_\ell} i \\
\varepsilon & \text{if } i = j \text{ and no self-loop} \\
\emptyset & \text{if no transition from } i \text{ to } j
\end{cases}
\]

\subsection*{Step 0: Base Cases $R^{(0)}_{ij}$}

Reading directly from the transition table (states $1 = p$, $2 = q$):

\begin{align*}
R^{(0)}_{11} &= a + \varepsilon = a^* \quad \text{(self-loop on } a \text{, plus } \varepsilon \text{ for staying)} \\[4pt]
R^{(0)}_{12} &= b \quad \text{(transition } p \xrightarrow{b} q \text{)} \\[4pt]
R^{(0)}_{21} &= a \quad \text{(transition } q \xrightarrow{a} p \text{)} \\[4pt]
R^{(0)}_{22} &= b + \varepsilon = b^* \quad \text{(self-loop on } b \text{, plus } \varepsilon\text{)}
\end{align*}

\medskip
\noindent\textit{Note:} Since $R^{(0)}_{ii}$ is used inside a Kleene star later, we keep them as $a+\varepsilon$ and $b+\varepsilon$ here for clarity, but note $(a+\varepsilon)^* = a^*$ and $(b+\varepsilon)^* = b^*$.

\subsection*{Step 1: $R^{(1)}_{ij}$ — Intermediate state allowed: $\{1\}$}

Using the recurrence with $k = 1$:
\[
R^{(1)}_{ij} = R^{(0)}_{i1}\left(R^{(0)}_{11}\right)^* R^{(0)}_{1j} + R^{(0)}_{ij}
\]

Note $(R^{(0)}_{11})^* = (a + \varepsilon)^* = a^*$.

\medskip
\noindent$\boldsymbol{R^{(1)}_{11}}$:
\[
R^{(1)}_{11} = R^{(0)}_{11}(R^{(0)}_{11})^* R^{(0)}_{11} + R^{(0)}_{11}
= (a+\varepsilon)\cdot a^* \cdot (a+\varepsilon) + (a+\varepsilon)
= a^* + (a+\varepsilon) = a^*
\]
\[
\boxed{R^{(1)}_{11} = a^*}
\]

\medskip
\noindent$\boldsymbol{R^{(1)}_{12}}$:
\[
R^{(1)}_{12} = R^{(0)}_{11}(R^{(0)}_{11})^* R^{(0)}_{12} + R^{(0)}_{12}
= (a+\varepsilon)\cdot a^* \cdot b + b
= a^*b + b = a^*b
\]
\[
\boxed{R^{(1)}_{12} = a^*b}
\]

\medskip
\noindent$\boldsymbol{R^{(1)}_{21}}$:
\[
R^{(1)}_{21} = R^{(0)}_{21}(R^{(0)}_{11})^* R^{(0)}_{11} + R^{(0)}_{21}
= a \cdot a^* \cdot (a+\varepsilon) + a
= a \cdot a^* + a = a^+
\]

Since $a \cdot a^* = a^+$ and $a^+ + a = a^+$:
\[
\boxed{R^{(1)}_{21} = a^+}
\]

\medskip
\noindent$\boldsymbol{R^{(1)}_{22}}$:
\[
R^{(1)}_{22} = R^{(0)}_{21}(R^{(0)}_{11})^* R^{(0)}_{12} + R^{(0)}_{22}
= a \cdot a^* \cdot b + (b+\varepsilon)
= a^+b + b + \varepsilon
= a^*b + \varepsilon
\]
\[
\boxed{R^{(1)}_{22} = a^*b + \varepsilon}
\]

\subsection*{Step 2: $R^{(2)}_{ij}$ — Intermediate states allowed: $\{1, 2\}$}

Using the recurrence with $k = 2$:
\[
R^{(2)}_{ij} = R^{(1)}_{i2}\left(R^{(1)}_{22}\right)^* R^{(1)}_{2j} + R^{(1)}_{ij}
\]

Note $(R^{(1)}_{22})^* = (a^*b + \varepsilon)^* = (a^*b)^*$.

\medskip
\noindent$\boldsymbol{R^{(2)}_{11}}$ \textbf{(needed for the final answer):}
\[
R^{(2)}_{11} = R^{(1)}_{12}(R^{(1)}_{22})^* R^{(1)}_{21} + R^{(1)}_{11}
= a^*b \cdot (a^*b)^* \cdot a^+ + a^*
= a^*b(a^*b)^* a^+ + a^*
= a^*(a^*b)^+ a^+ + a^*
\]

We can simplify $a^*b(a^*b)^*= a^*(ba^*)^*b$ (rewriting the iteration):
\[
\boxed{R^{(2)}_{11} = a^*\bigl(ba^+\bigr)^*}
\]

\textit{Verification of the simplification:} $a^*b(a^*b)^*a^+ + a^*$. Each term $b(a^*b)^*a^+$ represents one or more $b$-blocks each followed by one or more $a$'s. Factoring out the leading $a^*$:
\[
a^*b(a^*b)^* a^+ + a^* = a^*\bigl(b(a^*b)^* a^+ + \varepsilon\bigr) = a^*\bigl(ba^+)^* \text{ (by Arden's identity)}
\]

\medskip
\noindent\textit{The remaining expressions for completeness:}

\medskip
\noindent$\boldsymbol{R^{(2)}_{12}}$:
\[
R^{(2)}_{12} = R^{(1)}_{12}(R^{(1)}_{22})^* R^{(1)}_{22} + R^{(1)}_{12}
= a^*b(a^*b)^*(a^*b+\varepsilon) + a^*b
= a^*b(a^*b)^* + a^*b = a^*(ba^*)^*b
\]

\medskip
\noindent$\boldsymbol{R^{(2)}_{21}}$:
\[
R^{(2)}_{21} = R^{(1)}_{22}(R^{(1)}_{22})^* R^{(1)}_{21} + R^{(1)}_{21}
= (a^*b+\varepsilon)(a^*b)^* a^+ + a^+
= (a^*b)^* a^+ + a^+ = (a^*b)^* a^+
\]

\medskip
\noindent$\boldsymbol{R^{(2)}_{22}}$:
\[
R^{(2)}_{22} = R^{(1)}_{22}(R^{(1)}_{22})^* R^{(1)}_{22} + R^{(1)}_{22}
= (a^*b)^*(a^*b+\varepsilon) + (a^*b+\varepsilon)
= (a^*b)^* + \varepsilon = (a^*b)^*
\]

\section*{Resulting Regular Expression for $L(\mathcal{A})$}

The language $L(\mathcal{A})$ is accepted starting from the start state $p = 1$ and ending in any accept state. Since the only accept state is $p = 1$, we need all strings that start and end in state $1$, using any intermediate states $\{1,2\}$. This is given by $R^{(2)}_{11}$:

\[
\boxed{R \;=\; R^{(2)}_{11} \;=\; a^*\bigl(ba^+\bigr)^*}
\]

\subsection*{Interpretation}

The language $L(R) = L(\mathcal{A})$ consists of all strings over $\{a, b\}$ that:
\begin{itemize}
    \item start with zero or more $a$'s, then
    \item repeat zero or more times the pattern: one $b$ followed by one or more $a$'s.
\end{itemize}

In particular, every $b$ in an accepted string must be followed by at least one $a$. The string cannot end with $b$ (since $q$ is not an accept state). Examples: $\varepsilon,\ a,\ aa,\ aba,\ aaba,\ abba\ (\text{rejected}),\ bab\ (\text{rejected})$.

\subsubsection*{Given DFA (Figure 4.15)}

\begin{itemize}
    \item \textbf{States:} $A, B, C, D, E, F, G, H, I$
    \item \textbf{Start state:} $A$
    \item \textbf{Accept states:} $C, F, I$
\end{itemize}

\begin{center}
\begin{tabular}{c||c|c}
\toprule
\textbf{State} & \textbf{0} & \textbf{1} \\
\midrule
$\rightarrow A$ & $B$ & $E$ \\
$B$             & $C$ & $F$ \\
$*C$            & $D$ & $H$ \\
$D$             & $E$ & $H$ \\
$E$             & $F$ & $I$ \\
$*F$            & $G$ & $B$ \\
$G$             & $H$ & $B$ \\
$H$             & $I$ & $C$ \\
$*I$            & $A$ & $E$ \\
\bottomrule
\end{tabular}
\end{center}

\section*{Part (a): Table of Distinguishabilities}

\subsection*{Method}

Two states $p$ and $q$ are \textbf{distinguishable} if there exists some string $w$ such that exactly one of $\delta(p,w)$ and $\delta(q,w)$ is an accepting state. We build the table iteratively:

\begin{enumerate}
    \item \textbf{Base case:} Mark $\{p,q\}$ if exactly one is an accept state.
    \item \textbf{Propagation:} Mark $\{p,q\}$ if for some input $x \in \{0,1\}$, the pair $\{\delta(p,x),\,\delta(q,x)\}$ is already marked.
    \item Repeat until no new pairs can be marked.
\end{enumerate}

\subsection*{Base Case}

Accept states $= \{C, F, I\}$; \quad Non-accept states $= \{A, B, D, E, G, H\}$.

All pairs $\{p, q\}$ with one accept and one non-accept state are immediately marked distinguishable ($\times$).

\subsection*{Propagation}

We check all remaining pairs (both accept, or both non-accept):

\medskip
\noindent\textbf{Non-accept pairs that become distinguishable:}

\begin{align*}
\{A,B\} &: \delta(A,0)=B,\ \delta(B,0)=C \Rightarrow \{B,C\}\ \text{marked} \Rightarrow \textbf{marked}\ (\times) \\
\{A,E\} &: \delta(A,0)=B,\ \delta(E,0)=F \Rightarrow \{B,F\}\ \text{marked} \Rightarrow \textbf{marked}\ (\times) \\
\{A,H\} &: \delta(A,0)=B,\ \delta(H,0)=I \Rightarrow \{B,I\}\ \text{marked} \Rightarrow \textbf{marked}\ (\times) \\
\{B,D\} &: \delta(B,0)=C,\ \delta(D,0)=E \Rightarrow \{C,E\}\ \text{marked} \Rightarrow \textbf{marked}\ (\times) \\
\{B,G\} &: \delta(B,0)=C,\ \delta(G,0)=H \Rightarrow \{C,H\}\ \text{marked} \Rightarrow \textbf{marked}\ (\times) \\
\{D,E\} &: \delta(D,0)=E,\ \delta(E,0)=F \Rightarrow \{E,F\}\ \text{marked} \Rightarrow \textbf{marked}\ (\times) \\
\{D,H\} &: \delta(D,0)=E,\ \delta(H,0)=I \Rightarrow \{E,I\}\ \text{marked} \Rightarrow \textbf{marked}\ (\times) \\
\{E,G\} &: \delta(E,0)=F,\ \delta(G,0)=H \Rightarrow \{F,H\}\ \text{marked} \Rightarrow \textbf{marked}\ (\times) \\
\{G,H\} &: \delta(G,0)=H,\ \delta(H,0)=I \Rightarrow \{H,I\}\ \text{marked} \Rightarrow \textbf{marked}\ (\times)
\end{align*}

\medskip
\noindent\textbf{Pairs that remain indistinguishable (no marking possible):}

\begin{align*}
\{A,D\} &: \delta(\cdot,0)\to\{B,E\},\ \delta(\cdot,1)\to\{E,H\} \quad \text{(both unmarked)} \\
\{A,G\} &: \delta(\cdot,0)\to\{B,H\},\ \delta(\cdot,1)\to\{E,B\} \quad \text{(both unmarked)} \\
\{D,G\} &: \delta(\cdot,0)\to\{E,H\},\ \delta(\cdot,1)\to\{H,B\} \quad \text{(both unmarked)} \\
\{B,E\} &: \delta(\cdot,0)\to\{C,F\},\ \delta(\cdot,1)\to\{F,I\} \quad \text{(both accept pairs, unmarked)} \\
\{B,H\} &: \delta(\cdot,0)\to\{C,I\},\ \delta(\cdot,1)\to\{F,C\} \quad \text{(both accept pairs, unmarked)} \\
\{E,H\} &: \delta(\cdot,0)\to\{F,I\},\ \delta(\cdot,1)\to\{I,C\} \quad \text{(both accept pairs, unmarked)} \\
\{C,F\} &: \delta(\cdot,0)\to\{D,G\},\ \delta(\cdot,1)\to\{H,B\} \quad \text{(both unmarked)} \\
\{C,I\} &: \delta(\cdot,0)\to\{D,A\},\ \delta(\cdot,1)\to\{H,E\} \quad \text{(both unmarked)} \\
\{F,I\} &: \delta(\cdot,0)\to\{G,A\},\ \delta(\cdot,1)\to\{B,E\} \quad \text{(both unmarked)}
\end{align*}

\subsection*{Complete Distinguishability Table}

In the table below, $\times$ denotes a distinguishable pair and $\equiv$ denotes an indistinguishable (equivalent) pair. Pairs that are trivially marked (one accept, one non-accept) are shown as $\times$.

\begin{center}
\renewcommand{\arraystretch}{1.4}
\begin{tabular}{c|cccccccc}
  & $A$ & $B$ & $C$ & $D$ & $E$ & $F$ & $G$ & $H$ \\
\hline
$B$ & $\times$ & & & & & & & \\
$C$ & $\times$ & $\times$ & & & & & & \\
$D$ & $\equiv$ & $\times$ & $\times$ & & & & & \\
$E$ & $\times$ & $\equiv$ & $\times$ & $\times$ & & & & \\
$F$ & $\times$ & $\times$ & $\equiv$ & $\times$ & $\times$ & & & \\
$G$ & $\equiv$ & $\times$ & $\times$ & $\equiv$ & $\times$ & $\times$ & & \\
$H$ & $\times$ & $\equiv$ & $\times$ & $\times$ & $\equiv$ & $\times$ & $\times$ & \\
$I$ & $\times$ & $\times$ & $\equiv$ & $\times$ & $\times$ & $\equiv$ & $\times$ & $\times$ \\
\end{tabular}
\end{center}

\section*{Part (b): Minimum-State Equivalent DFA}

\subsection*{Equivalence Classes}

The indistinguishable pairs yield three equivalence classes:

\begin{center}
\begin{tabular}{cll}
\toprule
\textbf{New State} & \textbf{Original States} & \textbf{Role} \\
\midrule
$[A]$ & $\{A,\ D,\ G\}$ & Start state \\
$[B]$ & $\{B,\ E,\ H\}$ & Non-accept \\
$[C]$ & $\{C,\ F,\ I\}$ & Accept state \\
\bottomrule
\end{tabular}
\end{center}

\subsection*{Transition Table of Minimized DFA}

Using representative states to compute transitions:

\begin{center}
\begin{tabular}{c||c|c}
\toprule
\textbf{State} & \textbf{0} & \textbf{1} \\
\midrule
$\rightarrow [A]$ & $\delta(A,0)=B \to [B]$ & $\delta(A,1)=E \to [B]$ \\
$[B]$             & $\delta(B,0)=C \to [C]$ & $\delta(B,1)=F \to [C]$ \\
$*[C]$            & $\delta(C,0)=D \to [A]$ & $\delta(C,1)=H \to [B]$ \\
\bottomrule
\end{tabular}
\end{center}

\medskip
Written more compactly:

\begin{center}
\begin{tabular}{c||c|c}
\toprule
\textbf{State} & \textbf{0} & \textbf{1} \\
\midrule
$\rightarrow [A]$ & $[B]$ & $[B]$ \\
$[B]$             & $[C]$ & $[C]$ \\
$*[C]$            & $[A]$ & $[B]$ \\
\bottomrule
\end{tabular}
\end{center}

\subsection*{Diagram of Minimized DFA}

\begin{center}
\begin{tikzpicture}[
    shorten >=2pt,
    node distance=3cm,
    on grid,
    auto,
    >={Stealth[length=6pt]}
]
    \node[state, initial]           (A) {$[A]$};
    \node[state, right=of A]        (B) {$[B]$};
    \node[state, accepting, right=of B] (C) {$[C]$};

    \path[->]
        (A) edge node {0,1} (B)
        (B) edge node {0,1} (C)
        (C) edge [bend left=45] node {0} (A)
        (C) edge [bend right=35] node[below] {1} (B);
\end{tikzpicture}
\end{center}

\subsection*{Summary}

The original 9-state DFA is reduced to a \textbf{3-state} minimum DFA. The minimized automaton accepts strings whose length is congruent to $2 \pmod{3}$, i.e.\ strings of length $2, 5, 8, 11, \ldots$

\subsection*{Week 4 Discord Question}
When minimizing a DFA, does it matter which accepting state you use as a representative for an equivalence class when computing the transitions of the minimized DFA?

\subsection{Week 5}

\noindent Show that the following languages are not regular using the pumping lemma:
\begin{align*}
L_1 &= \{a^n b^m \in \{a,b\}^* \mid n, m \in \mathbb{N},\; n > m \geq 0\} \\
L_2 &= \{a^{n^2} \in \{a\}^* \mid n \in \mathbb{N},\; n \geq 1\} \\
L_3 &= \{a^p \in \{a\}^* \mid p \in \mathbb{N},\; p \text{ prime}\} \\
L_4 &= \{a^k b^m a^{k+m} \mid k, m \geq 0\}
\end{align*}
 
\section*{$L_1 = \{a^n b^m \mid n > m \geq 0\}$}
 
\textbf{Assume} $L_1$ is regular, so it satisfies the pumping lemma with constant $n$.
 
\medskip
\textbf{Choose} $w = a^{n+1}b^n \in L_1$ (since $n+1 > n \geq 0$), with $|w| = 2n+1 \geq n$. \checkmark
 
\medskip
\textbf{Decompose} $w = xyz$. By condition~2, $|xy| \leq n < n+1$, so $xy$ lies entirely within
the $a^{n+1}$ prefix. Hence $x = a^i$, $y = a^j$ with $j \geq 1$ (condition~1), and
$z = a^{n+1-i-j}b^n$.
 
\medskip
\textbf{Pump down} with $k = 0$:
\[
  w' = xz = a^{n+1-j}b^n.
\]
Since $j \geq 1$, we have $|w'|_a = n+1-j \leq n = |w'|_b$, so $w' \notin L_1$.
 
\medskip
\textbf{Contradiction.} $\square$
 
\section*{$L_2 = \{a^{n^2} \mid n \geq 1\}$}
 
\textbf{Assume} $L_2$ is regular with pumping constant $n$.
 
\medskip
\textbf{Choose} $w = a^{n^2} \in L_2$, with $|w| = n^2 \geq n$. \checkmark
 
\medskip
\textbf{Decompose} $w = xyz$. Since the alphabet is $\{a\}$, we have $x = a^i$, $y = a^j$,
$z = a^{n^2 - i - j}$ with $j \geq 1$ (condition~1) and $i + j \leq n$ (condition~2).
 
\medskip
\textbf{Pump} with $k = 2$:
\[
  w' = xy^2 z = a^{n^2 + j}.
\]
For $w' \in L_2$ we would need $n^2 + j = m^2$ for some integer $m \geq 1$. However,
since $1 \leq j \leq n$:
\[
  n^2 < n^2 + j \leq n^2 + n < n^2 + 2n + 1 = (n+1)^2.
\]
Thus $n^2 + j$ lies strictly between two consecutive perfect squares and is therefore not
a perfect square, giving $w' \notin L_2$.
 
\medskip
\textbf{Contradiction.} $\square$
 
\section*{$L_3 = \{a^p \mid p \text{ prime}\}$}
 
\textbf{Assume} $L_3$ is regular with pumping constant $n$.
 
\medskip
\textbf{Choose} a prime $p \geq n + 2$ (such a prime exists since there are infinitely many
primes), and let $w = a^p \in L_3$, so $|w| = p \geq n$. \checkmark
 
\medskip
\textbf{Decompose} $w = xyz$ with $x = a^i$, $y = a^j$ ($j \geq 1$, $i+j \leq n$),
$z = a^{p-i-j}$. Set $s := i + j$ so that $1 \leq s \leq n$.
 
\medskip
\textbf{Pump} with $k = p - j$. Note $k \geq 1$ because $j \leq n \leq p - 2 < p$.
The pumped word has length:
\[
  |xy^{p-j}z| = i + (p-j)\cdot j + (p - i - j)
               = p + j(p - j - 1)
               = p + pj - j^2 - j
               = p(1+j) - j(j+1)
               = (p - j)(1 + j).
\]
Since $p - j \geq p - n \geq 2$ and $j + 1 \geq 2$, the length $(p-j)(1+j)$ is a product of
two integers both at least $2$, hence \emph{composite}. Therefore $xy^{p-j}z \notin L_3$.
 
\medskip
\textbf{Contradiction.} $\square$
 
\section*{$L_4 = \{a^k b^m a^{k+m} \mid k, m \geq 0\}$}
 
\textbf{Assume} $L_4$ is regular with pumping constant $n$.
 
\medskip
\textbf{Choose} $w = a^n b^n a^{2n} \in L_4$ (setting $k = m = n$), with $|w| = 4n \geq n$. \checkmark
 
\medskip
\textbf{Decompose} $w = xyz$. By condition~2, $|xy| \leq n$, so $xy$ lies entirely within
the initial $a^n$ block. Hence $x = a^i$, $y = a^j$ with $j \geq 1$, and
$z = a^{n-i-j}b^n a^{2n}$.
 
\medskip
\textbf{Pump} with $k = 2$:
\[
  w' = xy^2 z = a^{n+j}\, b^n\, a^{2n}.
\]
For $w' \in L_4$ we would need $w' = a^{k'} b^{m'} a^{k'+m'}$ for some $k', m' \geq 0$.
Reading the structure of $w'$: $k' = n + j$ and $m' = n$, so the trailing $a$-block must
have length
\[
  k' + m' = (n + j) + n = 2n + j.
\]
But the trailing block of $w'$ has length $2n$. Since $j \geq 1$, we have $2n < 2n + j$,
so $w' \notin L_4$.
 
\medskip
\textbf{Contradiction.} $\square$

\subsection*{Week 5 Discord Question}
 Is it always possible to find such a "clever" k, or are there non-regular languages where no single pumping value works and you need a different proof strategy entirely?

\subsection{Week 6}
\section*{Exercise A Solutions}

\subsection*{1. TM for $L = \{1 0^n : n \in \mathbb{N}\}$ that outputs $1 0^{n+1}$}

\textbf{Idea:} Scan to the right over all symbols until the first blank is found, then write an additional $0$.

\textbf{Turing Machine:}
\[
M = (Q, \Sigma, \Gamma, \delta, q_0, B, F)
\]
where:
\begin{itemize}
    \item $Q = \{q_0, q_1, q_{accept}\}$
    \item $\Sigma = \{0,1\}$
    \item $\Gamma = \{0,1,B\}$
    \item $F = \{q_{accept}\}$
\end{itemize}

\textbf{Transition function:}
\[
\begin{aligned}
\delta(q_0, 1) &= (q_1, 1, R) \\
\delta(q_1, 0) &= (q_1, 0, R) \\
\delta(q_1, B) &= (q_{accept}, 0, R)
\end{aligned}
\]

\textbf{Example:} Input $100 \rightarrow$ Output $1000$

\bigskip

\subsection*{2. TM for $L = \{1 0^n : n \in \mathbb{N}\}$ that outputs $1$}

\textbf{Idea:} Keep the leading $1$ and erase all trailing $0$'s.

\textbf{Turing Machine:}
\[
M = (Q, \Sigma, \Gamma, \delta, q_0, B, F)
\]
where:
\begin{itemize}
    \item $Q = \{q_0, q_1, q_{accept}\}$
    \item $\Sigma = \{0,1\}$
    \item $\Gamma = \{0,1,B\}$
    \item $F = \{q_{accept}\}$
\end{itemize}

\textbf{Transition function:}
\[
\begin{aligned}
\delta(q_0, 1) &= (q_1, 1, R) \\
\delta(q_1, 0) &= (q_1, B, R) \\
\delta(q_1, B) &= (q_{accept}, B, R)
\end{aligned}
\]

\textbf{Example:} Input $1000 \rightarrow$ Output $1$

\bigskip

\subsection*{3. TM for all binary strings that swaps $0$ and $1$}

\textbf{Idea:} Traverse the tape and replace each $0$ with $1$ and each $1$ with $0$.

\textbf{Turing Machine:}
\[
M = (Q, \Sigma, \Gamma, \delta, q_0, B, F)
\]
where:
\begin{itemize}
    \item $Q = \{q_0, q_{accept}\}$
    \item $\Sigma = \{0,1\}$
    \item $\Gamma = \{0,1,B\}$
    \item $F = \{q_{accept}\}$
\end{itemize}

\textbf{Transition function:}
\[
\begin{aligned}
\delta(q_0, 0) &= (q_0, 1, R) \\
\delta(q_0, 1) &= (q_0, 0, R) \\
\delta(q_0, B) &= (q_{accept}, B, R)
\end{aligned}
\]

\textbf{Example:} Input $1010 \rightarrow$ Output $0101$

\section*{Exercise 1: Decidability and r.e. Languages}

We analyze each statement:

\bigskip

\textbf{1. If $L_1$ and $L_2$ are decidable, then so is $L_1 \cup L_2$.}

\textbf{True.}

\textbf{Argument:}  
Since $L_1$ and $L_2$ are decidable, there exist Turing machines $M_1$ and $M_2$ that halt on all inputs.  
Construct a machine $M$ that on input $w$:
\begin{itemize}
    \item Runs $M_1$ on $w$.
    \item Runs $M_2$ on $w$.
    \item Accepts if either accepts; otherwise rejects.
\end{itemize}
Since both machines halt, $M$ halts, so $L_1 \cup L_2$ is decidable.

\bigskip

\textbf{2. If $L$ is decidable, then so is $\overline{L} = \Sigma^* \setminus L$.}

\textbf{True.}

\textbf{Argument:}  
If $M$ decides $L$, construct $M'$ that:
\begin{itemize}
    \item Runs $M$ on input $w$.
    \item If $M$ accepts, reject.
    \item If $M$ rejects, accept.
\end{itemize}
Thus, $M'$ decides $\overline{L}$.

\bigskip

\textbf{3. If $L$ is decidable, then so is $L^*$.}

\textbf{True.}

\textbf{Argument:}  
Given input $w$, check all possible ways to split $w$ into substrings
\[
w = w_1 w_2 \cdots w_k
\]
and verify whether each $w_i \in L$. Since $L$ is decidable, each check halts.  
There are finitely many splits, so we can test all of them and decide whether $w \in L^*$.

\bigskip

\textbf{4. If $L_1$ and $L_2$ are r.e., then $L_1 \cup L_2$ is r.e.}

\textbf{True.}

\textbf{Argument:}  
Let $M_1$ and $M_2$ recognize $L_1$ and $L_2$. Construct a machine that:
\begin{itemize}
    \item Simulates both $M_1$ and $M_2$ in parallel (dovetailing).
    \item Accepts if either machine accepts.
\end{itemize}
Thus, $L_1 \cup L_2$ is recursively enumerable.

\bigskip

\textbf{5. If $L$ is r.e., then so is $\overline{L}$.}

\textbf{False.}

\textbf{Counterexample:}  
Let $L$ be the Halting Problem:
\[
L = \{ \langle M,w \rangle : M \text{ halts on input } w \}
\]
$L$ is r.e., but $\overline{L}$ (non-halting pairs) is not r.e.  
Thus, the statement is false.

\bigskip

\textbf{6. If $L$ is r.e., then so is $L^*$.}

\textbf{True.}

\textbf{Argument:}  
Given input $w$, enumerate all ways to split $w$ into substrings and simulate the recognizer for $L$ on each substring in parallel.  
If a decomposition exists where all substrings are accepted, then accept.  
Thus, $L^*$ is r.e.

\bigskip

\textbf{Summary:}
\begin{itemize}
    \item (1) True
    \item (2) True
    \item (3) True
    \item (4) True
    \item (5) False
    \item (6) True
\end{itemize}

\subsection*{Week 6 Discord Question}
Why does giving a Turing machine the ability to loop forever increase its computational power compared to DFAs?

\subsection{Week 7}

\section*{Problems on Decidability Exercise 2: Solution}

We classify each language as decidable, recursively enumerable (r.e.), or co-recursively enumerable (co-r.e.).

\subsection*{1. $L_1 = \{ M \mid M \text{ halts on itself} \}$}

\textbf{Claim:} $L_1$ is recursively enumerable but not decidable and not co-r.e.

\textbf{Proof:}
To show $L_1$ is recursively enumerable, we construct a Turing machine that simulates $M$ on input $M$. If $M$ halts on input $M$, the simulation halts and accepts. If $M$ does not halt, the simulation runs forever. Hence, $L_1$ is r.e.

To show $L_1$ is not decidable, note that this is a special case of the Halting Problem, which is undecidable.

To show $L_1$ is not co-r.e., observe that if both $L_1$ and its complement were recursively enumerable, then $L_1$ would be decidable, which is a contradiction.

\hfill $\square$

\subsection*{2. $L_2 = \{ (M,w) \mid M \text{ halts on } w \}$}

\textbf{Claim:} $L_2$ is recursively enumerable but not decidable and not co-r.e.

\textbf{Proof:}
To show $L_2$ is recursively enumerable, simulate $M$ on input $w$. If $M$ halts, accept; otherwise, run forever. Thus, $L_2$ is r.e.

$L_2$ is exactly the Halting Problem, which is known to be undecidable.

If $L_2$ were also co-r.e., then it would be decidable. Since it is not decidable, it cannot be co-r.e.

\hfill $\square$

\subsection*{3. $L_3 = \{ (M,w,k) \mid M \text{ halts on } w \text{ in at most } k \text{ steps} \}$}

\textbf{Claim:} $L_3$ is decidable (and hence both r.e.\ and co-r.e.).

\textbf{Proof:}
We construct a decider as follows: on input $(M,w,k)$, simulate $M$ on input $w$ for at most $k$ steps. If $M$ halts within $k$ steps, accept; otherwise, reject.

Since this procedure always halts after at most $k$ steps, $L_3$ is decidable. Therefore, it is also both recursively enumerable and co-recursively enumerable.

\hfill $\square$

\subsection*{Summary}

\begin{center}
\begin{tabular}{c|c|c|c}
Language & Decidable & r.e. & co-r.e. \\
\hline
$L_1$ & No & Yes & No \\
$L_2$ & No & Yes & No \\
$L_3$ & Yes & Yes & Yes \\
\end{tabular}
\end{center}

\section*{Landau Notation: Solutions to Exercises 2-5}

\subsection*{Exercise 2}

Let $f,g,h : \mathbb{N} \to \mathbb{R}_{\geq 0}$.

\begin{enumerate}
    \item \textbf{$f \in \mathcal{O}(f)$}

    Choose $M = 1$ and any $N \in \mathbb{N}$. Then for all $n \geq N$,
    \[
    f(n) \leq 1 \cdot f(n),
    \]
    hence $f \in \mathcal{O}(f)$.

    \item \textbf{If $c > 0$, then $\mathcal{O}(c f) = \mathcal{O}(f)$}

    ($\subseteq$) Suppose $h \in \mathcal{O}(c f)$. Then there exist $M,N$ such that
    \[
    h(n) \leq M \cdot c f(n) = (Mc) f(n),
    \]
    so $h \in \mathcal{O}(f)$.

    ($\supseteq$) Similarly, if $h \in \mathcal{O}(f)$, then
    \[
    h(n) \leq M f(n) = M \cdot \frac{1}{c} (c f(n)),
    \]
    so $h \in \mathcal{O}(c f)$.

    \item \textbf{If $f \leq g$ eventually, then $\mathcal{O}(f) \subseteq \mathcal{O}(g)$}

    If $f(n) \leq g(n)$ for all $n \geq N_0$, and $h \in \mathcal{O}(f)$, then
    \[
    h(n) \leq M f(n) \leq M g(n),
    \]
    for all sufficiently large $n$. Hence $h \in \mathcal{O}(g)$.

    \item \textbf{If $\mathcal{O}(f) \subseteq \mathcal{O}(g)$, then $\mathcal{O}(f+h) \subseteq \mathcal{O}(g+h)$}

    Let $k \in \mathcal{O}(f+h)$, so
    \[
    k(n) \leq M(f(n)+h(n)).
    \]
    Since $f \in \mathcal{O}(g)$, we have $f(n) \leq M' g(n)$ for large $n$. Thus,
    \[
    k(n) \leq M(M'g(n) + h(n)) \leq M''(g(n)+h(n)),
    \]
    hence $k \in \mathcal{O}(g+h)$.

    \item \textbf{If $h(n) > 0$ and $\mathcal{O}(f) \subseteq \mathcal{O}(g)$, then $\mathcal{O}(fh) \subseteq \mathcal{O}(gh)$}

    Let $k \in \mathcal{O}(fh)$. Then
    \[
    k(n) \leq M f(n)h(n).
    \]
    Since $f(n) \leq M' g(n)$ for large $n$,
    \[
    k(n) \leq MM' g(n)h(n),
    \]
    hence $k \in \mathcal{O}(gh)$.
\end{enumerate}

\subsection*{Exercise 3}

Let $i,j,k,n \in \mathbb{N}$.

\begin{enumerate}
    \item If $j \leq k$, then $n^j \leq n^k$ for $n \geq 1$, so
    \[
    \mathcal{O}(n^j) \subseteq \mathcal{O}(n^k).
    \]

    \item If $j \leq k$, then
    \[
    n^j + n^k \leq 2n^k,
    \]
    hence
    \[
    \mathcal{O}(n^j + n^k) \subseteq \mathcal{O}(n^k).
    \]

    \item For a polynomial,
    \[
    \sum_{i=0}^k a_i n^i \leq C n^k
    \]
    for some constant $C$, so
    \[
    \mathcal{O}\left(\sum_{i=0}^k a_i n^i\right) = \mathcal{O}(n^k).
    \]

    \item For $n \geq 2$, $\log n \leq n$, so
    \[
    \mathcal{O}(\log n) \subseteq \mathcal{O}(n).
    \]

    \item Since $\log n \leq n$,
    \[
    n \log n \leq n^2,
    \]
    hence
    \[
    \mathcal{O}(n \log n) \subseteq \mathcal{O}(n^2).
    \]
\end{enumerate}

\subsection*{Exercise 4}

\begin{enumerate}
    \item $\mathcal{O}(\sqrt{n}) \subseteq \mathcal{O}(n)$ since $\sqrt{n} \leq n$.

    \item $\mathcal{O}(n^2) \subseteq \mathcal{O}(2^n)$ since exponentials dominate polynomials.

    \item $\mathcal{O}(\log n) \subseteq \mathcal{O}((\log n)^2)$ since $\log n \leq (\log n)^2$ for large $n$.

    \item $\mathcal{O}(2^n) \subseteq \mathcal{O}(3^n)$ since $2^n \leq 3^n$.

    \item $\mathcal{O}(\log_2 n) = \mathcal{O}(\log_3 n)$ since logarithms differ by a constant factor:
    \[
    \log_2 n = \frac{\log_3 n}{\log_3 2}.
    \]
\end{enumerate}

\subsection*{Exercise 5}

We compare runtimes of sorting algorithms:

\begin{itemize}
    \item \textbf{Bubble Sort vs. Insertion Sort:}  
    Both have worst-case runtime $\mathcal{O}(n^2)$. However, insertion sort performs better on nearly sorted data ($\mathcal{O}(n)$ best case), while bubble sort generally remains inefficient.

    \item \textbf{Insertion Sort vs. Merge Sort:}  
    Insertion sort is $\mathcal{O}(n^2)$ in the worst case, while merge sort is $\mathcal{O}(n \log n)$ in all cases. Thus, merge sort is asymptotically more efficient.

    \item \textbf{Merge Sort vs. Quick Sort:}  
    Both have average runtime $\mathcal{O}(n \log n)$. However:
    \begin{itemize}
        \item Merge sort guarantees $\mathcal{O}(n \log n)$.
        \item Quick sort has worst-case $\mathcal{O}(n^2)$ but is often faster in practice due to smaller constants and cache efficiency.
    \end{itemize}
\end{itemize}

\textbf{Conclusion:}  
Asymptotically, algorithms with $\mathcal{O}(n \log n)$ dominate $\mathcal{O}(n^2)$ algorithms, while practical performance may still depend on constants and input characteristics.

\subsection*{Week 6 Discord Question}
Big O ignores constant factors and lower-order terms, but in real implementations those can still affect performance, so how do we decide when two algorithms in the same Big O class are actually meaningfully different in practice?

\subsection{Week 8}
\section*{Exercise 1}

\textbf{1.} 
\[
\varphi_1 := \neg((a \land b) \lor (\neg c \land d))
\]

\begin{align*}
\varphi_1 
&= \neg(a \land b) \land \neg(\neg c \land d) \\
&= (\neg a \lor \neg b) \land (c \lor \neg d)
\end{align*}

\textbf{CNF:}
\[
(\neg a \lor \neg b) \land (c \lor \neg d)
\]

\vspace{1em}

\textbf{2.}
\[
\varphi_2 := \neg((p \lor q) \rightarrow (r \land \neg s))
\]

\begin{align*}
\varphi_2 
&= \neg(\neg(p \lor q) \lor (r \land \neg s)) \\
&= \neg(\neg(p \lor q)) \land \neg(r \land \neg s) \\
&= (p \lor q) \land (\neg r \lor s)
\end{align*}

\textbf{CNF:}
\[
(p \lor q) \land (\neg r \lor s)
\]

\newpage

\section*{Exercise 2}

\textbf{1.}
\[
\psi_1 := (a \lor \neg b) \land (\neg a \lor b) \land (\neg a \lor \neg b)
\]

Try assignment: $a = \text{false},\ b = \text{false}$

\begin{align*}
(a \lor \neg b) &= F \lor T = T \\
(\neg a \lor b) &= T \lor F = T \\
(\neg a \lor \neg b) &= T \lor T = T
\end{align*}

\textbf{Conclusion:} Satisfiable.

\vspace{1em}

\textbf{2.}
\[
\psi_2 := (\neg p \lor q) \land (\neg q \lor r) \land \neg (\neg p \lor r)
\]

\begin{align*}
\neg(\neg p \lor r) &= p \land \neg r
\end{align*}

So:
\[
= (\neg p \lor q) \land (\neg q \lor r) \land p \land \neg r
\]

From $p \land \neg r$:
\[
p = T,\quad r = F
\]

Substitute:
\begin{align*}
(\neg p \lor q) &= F \lor q = q \\
(\neg q \lor r) &= \neg q \lor F = \neg q
\end{align*}

\[
q \land \neg q
\]

\textbf{Conclusion:} Not satisfiable.

\vspace{1em}

\textbf{3.}
\[
\psi_3 := (x \lor y) \land (\neg x \lor y) \land (x \lor \neg y) \land (\neg x \lor \neg y)
\]

From the clauses:
\[
(x \lor y) \land (\neg x \lor y) \Rightarrow y
\]
\[
(x \lor \neg y) \land (\neg x \lor \neg y) \Rightarrow \neg y
\]

\[
y \land \neg y
\]

\textbf{Conclusion:} Not satisfiable.

\subsection*{Week 8 Discord Question}
When simplifying a Boolean formula to check satisfiability, sometimes certain variable assignments seem “forced” by the clauses (like setting a variable to true or false to satisfy multiple clauses at once). Is there a systematic way to identify these forced assignments early, and how does this relate to algorithms like unit propagation?

\subsection{Week 9}

\section*{Exercise 1}

\subsection*{1a) Show that 
\[
\varphi=(a\lor b)\land(\neg a\lor c)\land(\neg b\lor c)\land(\neg c\lor d)\land \neg d
\]
is unsatisfiable using resolution.}

\subsubsection*{Step 1: Convert to clauses}

\[
\{\{a,b\},\{\neg a,c\},\{\neg b,c\},\{\neg c,d\},\{\neg d\}\}
\]

\subsubsection*{Step 2: Resolution Steps}

\[
\{\neg c,d\},\{\neg d\}\vdash \{\neg c\}
\]

\[
\{\neg a,c\},\{\neg c\}\vdash \{\neg a\}
\]

\[
\{\neg b,c\},\{\neg c\}\vdash \{\neg b\}
\]

\[
\{a,b\},\{\neg a\}\vdash \{b\}
\]

\[
\{b\},\{\neg b\}\vdash \Box
\]

\subsubsection*{Resolution Tree}

\begin{center}
\begin{tikzpicture}[node distance=1.8cm]
\node (ab) at (0,0) {$\{a,b\}$};
\node (ac) at (4,0) {$\{\neg a,c\}$};
\node (bc) at (8,0) {$\{\neg b,c\}$};
\node (cd) at (12,0) {$\{\neg c,d\}$};
\node (d) at (16,0) {$\{\neg d\}$};

\node (c) at (14,-2) {$\{\neg c\}$};
\node (na) at (6,-4) {$\{\neg a\}$};
\node (nb) at (10,-4) {$\{\neg b\}$};
\node (b) at (3,-6) {$\{b\}$};
\node (box) at (6,-8) {$\Box$};

\draw[->] (cd)--(c);
\draw[->] (d)--(c);
\draw[->] (ac)--(na);
\draw[->] (c)--(na);
\draw[->] (bc)--(nb);
\draw[->] (c)--(nb);
\draw[->] (ab)--(b);
\draw[->] (na)--(b);
\draw[->] (b)--(box);
\draw[->] (nb)--(box);
\end{tikzpicture}
\end{center}

Therefore, $\varphi$ is unsatisfiable.

\newpage

\subsection*{1b) Show that 
\[
\psi=p\land q\land(\neg p\lor \neg q\lor r)\land(\neg r\lor s)\land \neg s
\]
is unsatisfiable using unit resolution.}

\subsubsection*{Step 1: Convert to clauses}

\[
\{\{p\},\{q\},\{\neg p,\neg q,r\},\{\neg r,s\},\{\neg s\}\}
\]

\subsubsection*{Why is this Horn?}

Each clause contains at most one positive literal.

\subsubsection*{Unit Resolution Steps}

\[
\{p\},\{\neg p,\neg q,r\}\vdash \{\neg q,r\}
\]

\[
\{q\},\{\neg q,r\}\vdash \{r\}
\]

\[
\{r\},\{\neg r,s\}\vdash \{s\}
\]

\[
\{s\},\{\neg s\}\vdash \Box
\]

\subsubsection*{Resolution Tree}

\begin{center}
\begin{tikzpicture}[node distance=2cm]
\node (p) at (0,0) {$\{p\}$};
\node (q) at (3,0) {$\{q\}$};
\node (pqr) at (7,0) {$\{\neg p,\neg q,r\}$};
\node (rs) at (11,0) {$\{\neg r,s\}$};
\node (s) at (15,0) {$\{\neg s\}$};

\node (qr) at (5,-2) {$\{\neg q,r\}$};
\node (r) at (5,-4) {$\{r\}$};
\node (sp) at (8,-6) {$\{s\}$};
\node (box2) at (8,-8) {$\Box$};

\draw[->] (p)--(qr);
\draw[->] (pqr)--(qr);
\draw[->] (q)--(r);
\draw[->] (qr)--(r);
\draw[->] (r)--(sp);
\draw[->] (rs)--(sp);
\draw[->] (sp)--(box2);
\draw[->] (s)--(box2);
\end{tikzpicture}
\end{center}

Therefore, $\psi$ is unsatisfiable.

\newpage

\section*{Exercise 2: Drone Control System}

\subsection*{Step 1: Variables}

\begin{align*}
g &: \text{GPS lock valid} \\
b &: \text{battery healthy} \\
m &: \text{motor power enabled} \\
o &: \text{manual override present}
\end{align*}

\subsection*{Step 2: Formalization}

Policy:

\[
(g\land b)\to m
\]

\[
o\to m
\]

\[
\neg b \to \neg m
\]

Incident log:

\[
g,\quad \neg b,\quad o,\quad m
\]

Full formula:

\[
[(g\land b)\to m]\land(o\to m)\land(\neg b\to\neg m)\land g\land\neg b\land o\land m
\]

\subsection*{Step 3: Convert to Clauses}

\[
\{\neg g,\neg b,m\}
\]

\[
\{\neg o,m\}
\]

\[
\{b,\neg m\}
\]

\[
\{g\},\{\neg b\},\{o\},\{m\}
\]

So the clause set is:

\[
\{
\{\neg g,\neg b,m\},
\{\neg o,m\},
\{b,\neg m\},
\{g\},
\{\neg b\},
\{o\},
\{m\}
\}
\]

\subsection*{Step 4: Resolution}

\[
\{b,\neg m\},\{\neg b\}\vdash \{\neg m\}
\]

\[
\{\neg m\},\{m\}\vdash \Box
\]

\subsubsection*{Resolution Tree}

\begin{center}
\begin{tikzpicture}
\node (bm) at (0,0) {$\{b,\neg m\}$};
\node (nb) at (4,0) {$\{\neg b\}$};
\node (nm) at (2,-2) {$\{\neg m\}$};
\node (m) at (6,-2) {$\{m\}$};
\node (box3) at (4,-4) {$\Box$};

\draw[->] (bm)--(nm);
\draw[->] (nb)--(nm);
\draw[->] (nm)--(box3);
\draw[->] (m)--(box3);
\end{tikzpicture}
\end{center}

\subsection*{Conclusion}

The incident log is inconsistent with the policy because battery unhealthy implies motor must be disabled, but the log states motor power is enabled.

\[
m\land \neg m
\]

Hence, the system state is contradictory.

\subsection*{Week 9 Discord Question}
In Exercise 2, if the incident log did not explicitly state that motor power was enabled, would the remaining facts still conflict with the policy, or would the clauses become satisfiable?

\section{Synthesis}

\section{Evidence of Participation}

\section{Conclusion}\label{conclusion}

\begin{thebibliography}{99}
\bibitem[BLA]{bla} Author, \href{https://en.wikipedia.org/wiki/LaTeX}{Title}, Publisher, Year.
\end{thebibliography}

\end{document}
